Rough calculations

### 2015-11-01 notes/rough-calculations.Rmd
fixed bug in cline_from_soln

a.  Maximum cline width $\sigma \sqrt{\tau}$, even at unlinked loci.
b.  The selected cline takes time $s$ to spread out to its full width.
c.  Cline behaves like the selected cline for $r < 1/\tau$ (or, for $r < \sqrt{\frac{2}{s\tau}}$ if $\tau>1/s$).
d.  Cline is flattening for $x<\sigma\sqrt{\tau}$ and $r > \sqrt{\frac{2}{s\tau}}$.

@@

    But, what is the constant $C$?
    Consider the approximation of a lineage that acts like Brownian motion with a constant drift $s$ above zero.
    As noted in [this file](diffusion-local-times.html),
    the excursion measure of Brownian motion assigns mass $\sqrt{\frac{1}{2 \pi t^3}}$ to excursions of length $t$;
    we must tilt this by $e^{-ts}$ for the selection above zero;
    so that the mean rate at which this semi-pushed motion accumulates time above zero,
    with respect to local time at zero,
    is  (letting $x=\sqrt{2st}$)
    $$\begin{aligned}
    \int_0^\infty t \frac{e^{-st}}{ \sqrt{2 \pi t^3 } } dt 
    &=
    \int_0^\infty \frac{e^{-st}}{ \sqrt{2 \pi t } } dt \\
    &=
    \sqrt{2s} \int_0^\infty \frac{e^{-x^2/2}}{ \sqrt{2 \pi} } dx \\
    &=
    \sqrt{\frac{s}{2}} .
    \end{aligned}$$
    This suggests taking $C=\sqrt{s/2}$.

@@

a.  Look at space-time frequency images at $r=1$ and check they spread out as $\sqrt{t}$.

@@

c.  Look at space-recombination frequency images at a bunch of times
    to check that flattening occurs around $r=\sqrt{2/s\tau}$.

>>>>

a.  Maximum cline width $\sigma \sqrt{\tau}$, even at unlinked loci.
b.  The selected cline takes time $s$ to spread out to its full width.
c.  Cline behaves like the selected cline for $r < 1/\tau$ (or, for $r < \sqrt{\frac{2s}{\tau}}$ if $\tau>1/s$).
d.  Cline is flattening for $x<\sigma\sqrt{\tau}$ and $r > \sqrt{\frac{2s}{\tau}}$.

@@

    But, what is the constant $C$?
    Consider the approximation of a lineage that acts like Brownian motion with a constant drift $s$ above zero.
    As noted in [this file](diffusion-local-times.html),
    the excursion measure of Brownian motion assigns mass $\sqrt{\frac{1}{2 \pi t^3}}$ to excursions of length $t$;
    we must tilt this by $e^{-ts}$ for the selection above zero;
    so that the mean rate at which this semi-pushed motion accumulates time above zero,
    with respect to local time at zero,
    is  (letting $x=\sqrt{2st}$)
    $$\begin{aligned}
    \int_0^\infty t \frac{e^{-st}}{ \sqrt{2 \pi t^3 } } dt 
    &=
    \int_0^\infty \frac{e^{-st}}{ \sqrt{2 \pi t } } dt \\
    &=
    \sqrt{2s} \int_0^\infty \frac{e^{-x^2/2}}{ \sqrt{2 \pi} } dx \\
    &=
    \frac{1}{2\sqrt{2s}} .
    \end{aligned}$$
    This suggests taking $C=\sqrt{s/2}$.

@@

a.  Look at space-time frequency images at $r=1$ and check they spread out as $\sqrt{t}$.
    The green lines are at multiples of $\pm \sqrt{t}$.

@@

c.  Look at space-recombination frequency images at a bunch of times
    to check that flattening occurs around $r=\sqrt{2s/\tau}$.
    Note that the system is [stiff](https://en.wikipedia.org/wiki/Stiff_equation), 
    as diffusive systems [usually are](http://scicomp.stackexchange.com/questions/891/the-definition-of-stiff-ode-system),
    particularly for small $r$ at the times when each set of probabilities becomes steep;
    setting $s=0.1$ this is a problem, but not with smaller (less steep) $s$.

    ```{r space_recomb, cache=TRUE}
    s <- 0.01
    rvals <- exp(seq(log(1e-3),log(0.5),length.out=20))
    tt <- seq(0,200,length.out=10)
    thegrid <- setup.grid.1D(x.up=-100, x.down=100, N = 5000)
    fwds.soln <- forwards_pde(s=s,times=seq(0,1.2*max(tt),length.out=1000),grid=thegrid)
    r.solns <- lapply(rvals, function(r) {
            forwards_backwards_pde(s=s,times=tt,grid=thegrid,r=r,log.p=TRUE,fwds.soln=fwds.soln)
        } )
    unlinked.clines  <- lapply(r.solns, cline_from_soln, grid=thegrid )
    xr <- lapply( seq_along(tt), function (k) { do.call( cbind, lapply( unlinked.clines, function (x) x[k,-1] ) ) } )
    ```
    ```{r plot_sr}
    for (k in 2:length(xr)) {
        matplot(thegrid$x.mid, xr[[k]], type='l', col=rainbow(2*length(rvals)), lty=1, main=sprintf("t=%f",tt[k]) )
        legend("bottomright",lty=1, col=rainbow(2*length(rvals)), legend=sprintf("r=%f : %f",rvals,rvals*sqrt(tt[k]/2/s)))
    }
    ```
